---
title: "Interpolate to BOEM Regions"
output: html_document
date: "2022-09-13"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GoMex region

```{r}
librarian::shelf(
  arcpullr, dplyr, ggplot2, mapview)

dm_server <-"https://services2.arcgis.com/C8EMgrsFcRFL6LrL/ArcGIS/rest/services/"
lyr <- "DisMAP_Regions_20220516/FeatureServer/0"
lyr_url <- paste0(dm_server, lyr)
d_lyr <- get_spatial_layer(lyr_url) # table(d_lyr$OARegion)
table(d_lyr$OARegion)

regions <- c(
  gmx = "Gulf of Mexico",
  nef = "Northeast US Fall",
  nes = "Northeast US Spring",
  sef = "Southeast US Fall",
  ses = "Southeast US Spring",
  seu = "Southeast US Summer",
  wca = "West Coast Annual 2003-Present", # "West Coast Annual",
  wct = "West Coast Triennial 1977-2004") # West Coast Triennial")
regions_rds <- c(
  regions[!names(regions) %in% c("wca","wct")],
  wca = "West Coast Annual",
  wct = "West Coast Triennial")
# rgn    <- "gmx"
# region <- regions[[rgn]]

ply_rgns <- d_lyr %>% 
  group_by(OARegion) %>% 
  filter(OARegion %in% regions) %>% 
  summarize(
    geometry = st_union(geoms, is_coverage = T) %>% 
      st_cast("POLYGON")) %>% 
  st_as_sf(crs = 4326) %>% 
  mutate(
    one = 1) %>% 
  st_make_valid() %>% 
  st_simplify(dTolerance = 1) # 1 meter
mapview(ply_rgns)

ply_rgn <- ply_rgns %>% 
  filter(OARegion == region) %>% 

ply_rgn_mer <- st_transform(ply_rgn, 3857)
r_rgn_mer   <- raster(ply_rgn_mer, res=2000)
```

## Data

```{r}
# libraries ----
librarian::shelf(
  # fasterize, htmltools
  dplyr, glue, here, gstat, leaflet, mapview,
  purrr, raster, readr, sf, stringr, tibble)

select <- dplyr::select
options(readr.show_col_types = F)

# functions ----
basehere <- function(path){
  str_replace(path, paste0(here(), "/"), "")
}

get_sp_yr <- function(sp_key, yr){
  # sp_key="ach-spi1"; yr=2019
  
  r_yr_tif <- here(glue("tmp/{brgn}_{sp_key}_{yr}-1yr.tif"))
  
  if (!file.exists(r_yr_tif)){
    
    pts <- pts_rgn_mer %>% 
      filter(
        sp_key  == !!sp_key,
        year    == !!yr) 
    
    if (nrow(pts) == 0){
      message(glue("    unavailable: {brgn}_{sp_key} for {yr}"))
      return(NA)
    }
    
    if (nrow(pts %>% filter(wtcpue_cbrt > 0)) == 0){
      message(glue("    all zeros: {brgn}_{sp_key} for {yr}"))
      return(NA)
    }
    
    message(glue("    writing: {basehere(r_yr_tif)}"))
    mdl <- gstat(
      formula = wtcpue_cbrt ~ 1, locations = pts, 
      nmin = 10, nmax = 15, maxdist = 200000, set = list(idp = 1))
    r_yr <- interpolate(r_brgn_mer, mdl) %>% 
      mask(ply_brgn_mer) # mapview::mapview(r_idw)
    
    writeRaster(r_yr, r_yr_tif)
  }
  raster(r_yr_tif)
}

# paths ----
d_spp_csv     <- here("data_tif/_spp.csv")
d_spp_rgn_csv <- here("data_tif/_spp_rgn.csv")

# species list ----
if (any(!file.exists(d_spp_rgn_csv, d_spp_csv))){
  
  d_spp_rgn_0_csv <- here("data_clean/spplist.csv")
  
  d_spp_rgn <- read_csv(d_spp_rgn_0_csv) %>%
    rename(
      sp_sci = spp,
      sp_cmn = common)
   
  # * get unique species key `sp_key` ----
  sp_sci2key <- function(spp){
    gs <- str_split(spp, "\\W")[[1]][1:2]
    g <- str_sub(gs[1], end=3)
    s <- str_sub(gs[2], end=3)
    str_to_lower(glue("{g}-{s}"))
  }
  
  d_spp <- d_spp_rgn %>%
    group_by(sp_sci) %>%
    summarise(sp_cmn = first(sp_cmn)) %>%
    arrange(sp_sci, sp_cmn) %>%
    mutate(
      sp_key = map_chr(sp_sci, sp_sci2key))
  
  sp_keys_dup <- d_spp$sp_key[duplicated(d_spp$sp_key)]
  
  d_spp <- bind_rows(
    d_spp %>%
      filter(!sp_key %in% sp_keys_dup),
    d_spp %>%
      filter(sp_key %in% sp_keys_dup) %>%
      group_by(sp_key) %>%
      mutate(
        i = row_number(),
        sp_key2 = glue("{sp_key}{i}")) %>%
      select(
        sp_sci, sp_cmn, sp_key = sp_key2) ) %>%
    select(sp_key, sp_sci, sp_cmn) %>%
    arrange(sp_key, sp_sci)
  
  d_spp_rgn <- d_spp_rgn %>%
    left_join(
      d_spp %>%
        select(sp_sci, sp_key),
      by = "sp_sci")
    
  write_csv(d_spp, d_spp_csv)
  write_csv(d_spp_rgn, d_spp_rgn_csv)
}

d_spp     <- read_csv(d_spp_csv)
d_spp_rgn <- read_csv(d_spp_rgn_csv)

# BOEM regions ----

# list.files("data_clean", "dat_exploded.*rds$")

ply_brgns <- read_sf("~/Github/ecoquants/offhab-scripts/data/ply_rgns.geojson") # mapview(ply_brgns)
ply_brgns_mer <- st_transform(ply_brgns, 3857)

bregions <- c(
  waor = "Washington/Oregon",       # light green
  scal = "Southern California",     # dark green
  natl = "North Atlantic",          # light blue
  matl = "Mid Atlantic",            # dark blue
  cgmx = "Central Gulf of Mexico")  # pink

brgn_pal <- RColorBrewer::brewer.pal(
  length(bregions), "Paired")
names(brgn_pal) = names(bregions)

brgn2rgns <- list(
  cgmx = c("gmx"),
  waor = c("wca"),
  scal = c("wca"),
  natl = c("nef","nes"),
  matl = c("nef","nes","sef","ses","seu"))

# iterate over BOEM regions ----
for (brgn in names(bregions)){ # brgn = names(bregions)[1] # brgn = "cgmx"
  bregion <- bregions[[brgn]]
  message(glue("BOEM REGION: {brgn} ({bregion})"))
  
  # get data from all regions
  stopifnot(brgn %in% names(brgn2rgns))
  rgns <- brgn2rgns[[brgn]]
  d_lst <- lapply(
    rgns, function(rgn){
      # browser()
      region_rds <- regions_rds[[rgn]]
      d <- readRDS(glue("data_clean/dat_exploded{region_rds}.rds")) })
  d_rgn <- bind_rows(d_lst)

  # rgn    <- "gmx"
  # region <- regions[[rgn]]
  
  pts_rgn <- d_rgn %>% 
    # tibble() %>% 
    mutate(
      year        = as.integer(year),
      wtcpue_cbrt = wtcpue^(1/3)) %>% # take the cube root
    rename(
      sp_sci = spp) %>%
    left_join(
      d_spp %>%
        select(sp_sci, sp_key),
      by = "sp_sci") %>%
    st_as_sf(
      coords = c("lon", "lat"), remove = T, crs = 4326) %>%
    select(
      year, sp_key, wtcpue_cbrt)  %>% 
    filter(!is.na(sp_key)) # gmx nrow: 12,921,639 -> 2,525,473
   # dropping: region, haulid, common, stratum, stratumarea, depth, wtcpue
  pts_rgn_mer <- st_transform(pts_rgn, 3857)
  
  # brgn <- "cgmx"
  # bregion <- bregions[[brgn]]
  ply_brgn_mer <- ply_brgns_mer %>%
    filter(RESA_summa == bregion)
  r_brgn_mer <- raster(ply_brgn_mer, res=2000)
  
  # iterate over species ----
  if (!all(regions[rgns] %in% unique(d_spp_rgn$region))){
    message("WHOAH! why are rgns not found?!")
  }
  sp_keys <- d_spp_rgn %>% 
    filter(
      !flagged,
      sp_key != "na-na",
      region %in% regions[rgns]) %>% 
    distinct(sp_key) %>% 
    arrange(sp_key) %>% 
    pull(sp_key)
  
  for (sp_key in sp_keys){ # sp_key = sp_keys[1]
    
    d_sp <- d_spp_rgn %>% 
      filter(
        region == !!region, 
        sp_key == !!sp_key)
    message(glue("SPECIES: {sp_key} ({d_sp$sp_sci}) {d_sp$sp_cmn}"))
    
    yrs <- pts_rgn_mer %>% 
      st_drop_geometry() %>% 
      tibble() %>% 
      filter(sp_key == !!sp_key) %>% 
      arrange(year) %>% 
      pull(year) %>% 
      unique()
    if (length(yrs) == 0){
      message("  WHOAH! why zero yrs?! SKIPPING")
      next()
    }
    
    #for (yr in yrs){ # yr = yrs[1]
    for (yr in max(yrs)){ # yr =  max(yrs)
      
      message(glue("  YEAR: {yr}"))
  
      r_yrs_tif <- here(glue("data_tif/{brgn}_{sp_key}_{yr}.tif"))
      
      if (!file.exists(r_yrs_tif)){
        message(glue("    TRYING: {basehere(r_yrs_tif)}"))
        
        yrs <- (yr-2):(yr+2)
        wts <- 3 - (abs(yr - yrs)) # 1 2 3 2 1
        lst <- sapply(
          yrs, function(yr){
            get_sp_yr(sp_key, yr) })
        idx <- !is.na(lst)
        if (sum(idx) == 0){
          message("    SKIPPING: no data")
          next()
        }
        message(glue("    WRITING: {basehere(r_yrs_tif)}"))
        stk <- stack(lst[idx])
        r_yrs <- raster::weighted.mean(stk, wts[idx], na.rm=F)
        # mapview(r_yrs)
        writeRaster(r_yrs, r_yrs_tif)
      }
    }
  }
}
```

```{r}
# plot(r_idw)
# r_idw <- r_idw^3

# Create a continuous palette function
pal <- colorNumeric(
  palette = "viridis",
  domain = values(r_idw),
  na.color = "transparent", alpha = T)

leaflet() %>%
  addProviderTiles(providers$Esri.OceanBasemap) %>% 
  addRasterImage(
    r_idw, colors = pal, project = F) %>%
  addLegend(
    pal = pal,
    values = values(r_idw),
    title = paste0(
      sp_common, 
      "<br>∛wtcpue<br>",
      min(yrs), ":", max(yrs)),
    opacity = 1)
```


The predicted biomass density (kg per tow) distribution
based on fishery-independent survey data. The
distribution surface is created by applying the inverse
distance weighting (IDW) interpolation algorithm to the
observations in the survey for each species, regions, and
season combination. The grid size is 2km x 2 km. The IDW
approach smooths over multiple observations to
interpolate the biomass across areas where the survey did
not sample. For more detailed information on the
methodology and data sources see the DisMAP 


```{r}


url <- "https://maps.fisheries.noaa.gov/image/rest/services/DisMAP/Gulf_of_Mexico_20220516/ImageServer"

crs_txt <- 'PROJCS["WGS_1984_Albers_NMSDD",GEOGCS["GCS_WGS_1984",DATUM["D_WGS_1984",SPHEROID["WGS_1984",6378137.0,298.257223563]],PRIMEM["Greenwich",0.0],UNIT["Degree",0.0174532925199433]],PROJECTION["Albers"],PARAMETER["false_easting",0.0],PARAMETER["false_northing",0.0],PARAMETER["central_meridian",180.0],PARAMETER["standard_parallel_1",-2.0],PARAMETER["standard_parallel_2",49.0],PARAMETER["latitude_of_origin",25.5],UNIT["Meter",1.0]]'
crs_alb <- st_crs(crs_txt)

ply_gm_alb <- st_transform(ply_gm, crs_alb)

class(ply_gm_alb)

bb <- paste(sf::st_bbox(ply_gm_alb), collapse =",")
img <- get_image_layer(url, bbox = bb)

arcpullr:::get_raster_layer
img

response_raw <- httr::POST(url = url, body = list(f = "json", 
        token = "", bbox = bb, bboxSR = bbox_sr, 
        imageSR = bbox_sr, transparent = transparent, format = format, 
        ...))

sf_obj <- ply_gm_alb
bbox <- v
bbox_coords <- paste(bbox, collapse = ", ")
bbox_sr <- get_sf_crs(sf_obj)
sf::st_crs(sf_obj)[[1]]

plot_layer(img)




```

